---
title: "AdvertisingResponse_RMD"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## First Steps 

NOTE: Your data needs to be in the folder where your code is located!

The first Chunk of Code will always do 3 Things for us

1) Clear ALL Variables (to make sure nothing is "left over") and clear the      screen. 
2) Read in the data we want to use. In our case the file is called              "Data_Advertising_R.csv)"
3) Explore the data (what variables are in it and what are some basic           statistics for these variables)

```{r}

# Clear All Variables & Clear the Screen
rm(list=ls())
cat("\014")

# Read in the Data
data.adv = read.csv("Data_Advertising_R.csv")

# Explore the data
str(data.adv)
summary(data.adv)
```


## PREPARE DATA
 
In our case we have to create new variables: log(Advertising), Advertising^2, and Lag(Sales)

```{r}

# Create log Advertising
data.adv$logAdvertising <- log(data.adv$Advertising)

# Create Advertising^2
data.adv$Advertising2 <- (data.adv$Advertising)^2

# Create LAG Sales
data.adv$lagSales <- c(NA, head(data.adv$Sales, -1))

# Display the data
summary(data.adv)
```


## REGRESSION MODELS

1) Define y (Dependent Variable). For this example, Sales is used.

2) Define X (Independent Variables). For this example, Advertising and Lagged    Sales (i.e., a model with carry-over) are used

3) Run the Models. For example, a linear model where Sales (Sales) are a        function of Advertising (Advertising) and lagged Sales (lagSales)

```{r}

# Run the Regression (includes an INTERCEPT) - The Models Numbers are the ones from the deck
lm.model1 <- lm(Sales ~ Advertising, data = data.adv)
lm.model2 <- lm(Sales ~ logAdvertising, data = data.adv)
lm.model3 <- lm(Sales ~ Advertising + Advertising2, data = data.adv)

lm.model4 <- lm(Sales ~ Advertising + lagSales, data = data.adv)
lm.model5 <- lm(Sales ~ logAdvertising + lagSales, data = data.adv)
lm.model6 <- lm(Sales ~ Advertising + Advertising2 + lagSsales, data = data.adv)

# Display Results
summary(lm.model1)
summary(lm.model2)
summary(lm.model3)
summary(lm.model4)
summary(lm.model5)
summary(lm.model6)
```
```

