---
title: "Price Response"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## First Steps 

NOTE: Your data needs to be in the folder where your code is located!

The first Chunk of Code will always do 3 Things for us

1) Clear ALL Variables (to make sure nothing is "left over") and clear the      screen. 
2) Read in the data we want to use. In our case the file is called              "data.price.csv)"
3) Explore the data (what variables are in it and what are some basic           statistics for these variables)

```{r}

# Clear All Variables & Clear the Screen
rm(list=ls())
cat("\014")

# Read in the Data
data.price = read.csv("Data_Price_R.csv")

# Explore the data
str(data.price)
summary(data.price)
```


## PREPARE DATA
 
In our case we have to create new variables: log(Units) and log(Prices)

```{r}

# Create log UNITS
data.price$logUNITS1 <- log(data.price$UNITS1)
data.price$logUNITS2 <- log(data.price$UNITS2)
data.price$logUNITS3 <- log(data.price$UNITS3)
data.price$logUNITS4 <- log(data.price$UNITS4)

# Create log PRICES
data.price$logPRICE1 <- log(data.price$PRICE1)
data.price$logPRICE2 <- log(data.price$PRICE2)
data.price$logPRICE3 <- log(data.price$PRICE3)
data.price$logPRICE4 <- log(data.price$PRICE4)
```


## REGRESSION MODELS

1) Define y (Dependent Variable). For this example, Sales for Brand 1           (UNITS1) is used.

2) Define X (Independent Variables). For this example, OWN Marketing Activity    (PRICE1, FEAT1 and DISP1) is used

3) Run the Models. For example, a linear model where Sales for Brand 1          (UNITS1) are a function of Units1's Price (PRICE1), Feature (FEAT1), and     Display (DISP1)

```{r}

# Run the Regression (includes an INTERCEPT)

# Models from Slide "PRICE REGRESSIONS – EXAMPLE OREO"
# These are the basic Models that only have the Price of Oreo in them
lm.linear   <- lm(UNITS1 ~ PRICE1, data = data.price)
lm.semi_log <- lm(logUNITS1 ~ PRICE1, data = data.price)
lm.log_log  <- lm(logUNITS1 ~ logPRICE1, data = data.price)


# Models from Slide "INCORPORATING PROMOTIONAL RESPONSE", results are on the next slide in the deck
# these models have next to Oreo's own price also Feature and Display for Oreo
# Note we cannot run the log-log model as Feature and Display include zeros so we only run two models
lm.linear_own   <- lm(UNITS1 ~ PRICE1+FEAT1+DISP1, data = data.price)
lm.semi_log_own <- lm(logUNITS1 ~ PRICE1+FEAT1+DISP1, data = data.price)

# Models from Slide "CROSS-PRICE EFFECTS AND CROSS-PROMOTION EFFECTS", results are on the next slide in the deck
# these models have next to Oreo's own price also Feature and Display for Oreo as well as ALL competitive marketing activity
# Note we cannot run the log-log model as Feature and Display include zeros so we only run two models
# Note that there are no Display and Feature for Brand 3 and no Feature for Brand 4
lm.linear_own_comp   <- lm(UNITS1 ~ PRICE1+FEAT1+DISP1+PRICE2+FEAT2+DISP2 +PRICE3+PRICE4+FEAT4, data = data.price)
lm.semi_log_own_comp <- lm(logUNITS1 ~ PRICE1+FEAT1+DISP1+PRICE2+FEAT2 +DISP2+PRICE3+PRICE4+FEAT4, data = data.price)

# Display Results
summary(lm.linear)
summary(lm.semi_log)
summary(lm.linear_own)
summary(lm.semi_log_own)
summary(lm.linear_own_comp)
summary(lm.semi_log_own_comp)

# How caculate a Price Elasticity - example Linear Model
# Use the command str(lm.linear) to see what's in the data frame lm.linear
price.coeff.linear = unname(lm.linear$coefficients[2], force = FALSE)
elas.linear        = price.coeff.linear*mean(data.price$PRICE1)/mean(data.price$UNITS1)

```
