---
title: 'Assignment 3'
author: "Team 7 - Aakash and Noufris"
date: "May 2023"
output: word_document
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Clean the environment and import libraries
```{r}
rm(list = ls())
cat('\014')
library(broom)
library(knitr)
library(psych)
library(dplyr)
library(ggplot2)
```

### Question 1: Data Description
Loading data

```{r}
load ('C:\\Users\\Aakash\\Desktop\\UW Courses\\Codes_MSBA\\Analytics for firm decisions\\Assignment 3\\Detergent.RData')

```
Describing data

```{r}
describe(detergent_DF)

```
Creating total sales per week for each detergent

```{r}
detergent_DF$weekly_sales_tide128 <- detergent_DF$q_tide128*detergent_DF$p_tide128
detergent_DF$weekly_sales_tide64 <- detergent_DF$q_tide64*detergent_DF$p_tide64
detergent_DF$weekly_sales_wisk64 <- detergent_DF$q_wisk64*detergent_DF$p_wisk64

total_sales <- sum(c(detergent_DF$weekly_sales_tide128, detergent_DF$weekly_sales_tide64, detergent_DF$weekly_sales_wisk64))

marketshare_tide128 <- sum(detergent_DF$weekly_sales_tide128)/total_sales
marketshare_tide64 <- sum(detergent_DF$weekly_sales_tide64)/total_sales
marketshare_wisk64 <- sum(detergent_DF$weekly_sales_wisk64)/total_sales

mean_price_tide128 <- mean(detergent_DF$p_tide128)
median_price_tide128 <- median(detergent_DF$p_tide128)
mean_price_tide64 <- mean(detergent_DF$p_tide64)
median_price_tide64 <- median(detergent_DF$p_tide64)
mean_price_wisk64 <- mean(detergent_DF$p_wisk64)
median_price_wisk64 <- median(detergent_DF$p_wisk64)
sd_price_tide128 <- sd(detergent_DF$p_tide128)
sd_price_tide64 <- sd(detergent_DF$p_tide64)
sd_price_wisk64 <- sd(detergent_DF$p_wisk64)
```


: Table of market share (in percentage) and price statistics (in dollars)

+---------------+---------------+-------------+--------------+-----------+
| Product       | Market share  | Mean Price  | Median Price | Std. Dev  |          
+===============+===============+=============+==============+===========+
| Tide 128 oz   |   56.8%       |   8.363     |   8.476      |  0.760    |
+---------------+---------------+-------------+--------------+-----------+
| Tide 64 oz    |   26.3%       |   4.375     |   4.419      |  0.404    |
+---------------+---------------+-------------+--------------+-----------+
| Wisk 64 oz    |   16.8%       |   4.072     |   4.190      |  0.490    |
+---------------+---------------+-------------+--------------+-----------+


Creating the price gaps between products
```{r}
detergent_DF$p_gap_tide128_64 <- detergent_DF$p_tide128-detergent_DF$p_tide64
detergent_DF$p_gap_tide64_wisk64 <- detergent_DF$p_tide64-detergent_DF$p_wisk64

mean_p_gap_tide128_64 <- mean(detergent_DF$p_gap_tide128_64)
median_p_gap_tide128_64 <- median(detergent_DF$p_gap_tide128_64)
sd_p_gap_tide128_64 <- sd(detergent_DF$p_gap_tide128_64)

mean_p_gap_tide64_wisk64 <- mean(detergent_DF$p_gap_tide64_wisk64)
median_p_gap_tide64_wisk64 <- median(detergent_DF$p_gap_tide64_wisk64)
sd_p_gap_tide64_wisk64 <- sd(detergent_DF$p_gap_tide64_wisk64)

```

: Table of the mean, median, and std. dev. of the two price gap variables

+----------------------------+-------------+-------------+--------------+
| Gap                        |    Mean     |   Median    |  Std. Dev    |        
+============================+=============+=============+==============+
| Tide 128 oz & Tide 64 oz   |    3.988    |   4.094     |   0.870      |
+----------------------------+-------------+-------------+--------------+
| Tide 64 oz & Wisk 64 oz    |    0.303    |   0.261     |   0.586      |
+----------------------------+-------------+-------------+--------------+

Histograms of the price gaps
```{r}


ggplot(detergent_DF, aes(x = p_gap_tide128_64)) +
  geom_histogram(binwidth = 1, fill = "darkmagenta", color = "black") +
  stat_bin(binwidth = 1, geom = "text", aes(label = after_stat(count), y = after_stat(count)), vjust = -0.5, size = 3) +
  labs(x = "Price Gap ($)", y = "Frequency", title = "Price Gap between Tide 128 oz & Tide 64 oz") +
  theme_minimal() +
  theme(panel.grid.major = element_line(colour = "gray", linetype = "dotted"),
        plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
            axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15))


ggplot(detergent_DF, aes(x = p_gap_tide64_wisk64)) +
  geom_histogram(binwidth = 1, fill = "cyan2", color = "black") +
  stat_bin(binwidth = 1, geom = "text", aes(label = after_stat(count), y = after_stat(count)), vjust = -0.5, size = 3) +
  labs(x = "Price Gap ($)", y = "Frequency", title = "Price Gap between Tide 64 oz & Wisk 64 oz") +
  theme_minimal() +
  theme(panel.grid.major = element_line(colour = "gray", linetype = "dotted"),
        plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
            axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15))

```

The distribution of the price variation between Tide 128oz and Tide 64oz is approximately a normal distribution. The coefficient of variation is ~20% indicating that relatively the deviation is low and most prices are distributed around mean price of roughly 4. The range of the distribution is from -0.1 to 7.2.

The distribution of the price variation between Tide 64oz and Whisk 64oz is a right skewed distribution. The coefficient of variation is ~190% indicating that relatively the deviation is high and most prices are distributed to the left of mean price of roughly 0.3. The range of the distribution is from -3.52 to 2.96.


### Question 2: Demand Estimation
Construct the sales velocity for each of Tide 64 and Tide 128

```{r}
detergent_DF$velocity_tide128 <- detergent_DF$q_tide128/detergent_DF$acv
detergent_DF$velocity_tide64 <- detergent_DF$q_tide64/detergent_DF$acv
```

All Commodity Volume (ACV) measures the sales potential of a specific product or category within a store or market. Dividing unit sales by the ACV gives us sales velocity which measures the product's performance and helps us to evaluate how well a product is selling relative to the total sales potential of the store. 

The purpose of diving the unit sales by ACV is to capture the impact of price on the demand in sales volume effectively. By dividing sales with the sales potential of the store, we are making the sales made across diffrent stores comparable and removing the baises it might have in terms of potential of making those sales.

Now, we make log-linear demand models for the two Tide products by regressing the log of velocity on all prices (own and competing products)
```{r}
reg_tide128 <- lm(formula = log(velocity_tide128) ~ log(p_tide128) + log(p_tide64) + log(p_wisk64), 
                  data = detergent_DF)
reg_tide64 <- lm(formula = log(velocity_tide64) ~ log(p_tide128) + log(p_tide64) + log(p_wisk64), 
                  data = detergent_DF)

summary(reg_tide128)
summary(reg_tide64)
```
According to the regression results of Tide 128 oz, we observe that the product is highly elastic to its own price changes having an own price elasticity of -4.5971. 
In addition, a price increase of Tide 64 oz will lead to increased demand of Tide 128 oz (cross price elasticity of 0.2867). Moreover, a price increase of Wisk 64 oz will lead to increased demand of Tide 128 oz (cross price elasticity of 0.1514).
We observe that a price change of Tide 64 oz has more effect on the demand of Tide 128 oz than a price change of Wisk 64 oz.

According to the regression results of Tide 64 oz, we observe that the product is highly elastic to its own price changes having an own price elasticity of -3.7486.
In addition, a price increase of Tide 128 oz will lead to increased demand of Tide 64 oz (cross price elasticity of 1.4479). Moreover, a price increase of Wisk 64 oz will lead to decreased demand of Tide 64 oz (cross price elasticity of -0.8755).
We observe that a price change of Tide 128 oz has a positive effect on the demand of Tide 64 oz, whereas a price change of Wisk 64 has a negative effect on the demand of Tide 64 oz. The reason of why price increase in Wisk 64oz leads to lower sales in Tide 64oz, needs to be investigated by looking into broader aspects at customer, product and market level. Maybe there is a confounding factor at play in here.  

### Question 3: Time Trend
Re-estimate the log-linear demand models for the two Tide products including a time trend

```{r}
reg_tide128_time <- lm(formula = log(velocity_tide128) ~ log(p_tide128) + log(p_tide64) + 
                         log(p_wisk64) + factor(week), data = detergent_DF)
reg_tide64_time <- lm(formula = log(velocity_tide64) ~ log(p_tide128) + log(p_tide64) + 
                         log(p_wisk64) + factor(week), data = detergent_DF)

kable((tidy(reg_tide128_time))[1:4,], digits = 4)
kable((tidy(reg_tide64_time))[1:4,], digits = 4)
```

The time trend is important to assess the seasonality which might impact our analysis. After adding in the week number as a time trend distinction, we get the following results:

For Tide 128 oz:

We observe that the own price elasticity drops to -3.1934. 
In addition, the cross price elasticity of Tide 64 oz on Tide 128 oz jumps up to 0.5950.
Finally, the cross price elasticity of Wisk 64 oz on Tide 128 oz becomes -0.3468, which means that in this model, increase in price of Wisk 64 oz affects the sales of 128 oz Tide negatively. This is again interesting and the reason for this needs to be further investigated.

For Tide 64 oz:

We observe that there is a negative affect on sales by increase in price of either of Tide 128 oz, Tide 64 oz or Wisk 64 Oz, compared to earlier model of negative affect on only own price and the price of Wisk 64 oz. This is again interesting and the reason for this needs to be further investigated.

For both of the new models with time trend included, there seems to be a confounder affecting the sales of the products such as ongoing promotions, product placement etc.

### Question 4: Focus on Non-Promoted Weeks
Fraction of store-weeks where at least one of the detergents was promoted

```{r}
mean(detergent_DF$promoflag)
```
Create a new dataframe including the non-promoted store-weeks

```{r}
detergent_DF_2 = subset(detergent_DF, promoflag != 1)
```
Re-estimate the log-linear demand models with a time-trend for the two Tide products only using
data from non-promoted store-weeks

```{r}
reg_tide128_time_nonpromoted <- lm(formula = log(velocity_tide128) ~ log(p_tide128) + log(p_tide64) + 
                         log(p_wisk64) + factor(week), data = detergent_DF_2)
reg_tide64_time_nonpromoted <- lm(formula = log(velocity_tide64) ~ log(p_tide128) + log(p_tide64) + 
                         log(p_wisk64) + factor(week), data = detergent_DF_2)

kable((tidy(reg_tide128_time_nonpromoted))[1:4,], digits = 4)
kable((tidy(reg_tide64_time_nonpromoted))[1:4,], digits = 4)
```

According to the regression results of Tide 128 oz without promotions, we observe the following:
1. The magnitude of own price elasticity decreases in magnitude, being -3.11. 
2. The magnitude of cross price elasticity of Tide 64 oz on Tide 128 oz increases in magnitude to 1.0261 
3. The magnitude of the cross price elasticity of Wisk 64 oz on Tide 128 oz increases in magnitude to -1.0796 and having a negative effect on the demand.

Based on the regression results, this model makes more sense for the following reasons:
1. Without having any promotions, the magnitude of the own price elasticity is less excessive. 
2. A price change of Tide 64 oz has a significantly greater impact on the demand of Tide 128 oz than the previous model. This indicates a strong cannibalization of the sales.

According to the regression results of Tide 64 oz without promotions, we observe the following:
1. The magnitude of own price elasticity decreases in magnitude to -0.6663. 
2. The magnitude of cross price elasticity of Tide 128 oz on Tide 64 oz increases in magnitude significantly to -1.5159 and having a greater negative effect on the demand.
3. The magnitude of the cross price elasticity of Wisk 64 oz on Tide 64 oz increases in magnitude to -0.6066, having more negative effect on the demand.

Based on the regression results, this model makes more sense for the following reasons:
1. Without having any promotions, the magnitude of the own price elasticity is less excessive. 
2. A price change of Tide 128 oz has a greater effect on the demand of Tide 64 oz than the previous model.

### Question 5: Store Fixed Effects
Re-estimate the log-linear demand models for the two Tide products including a time trend and store
fixed effects using the data for the non-promoted store-weeks
```{r}
reg_tide128_time_nonpromoted_fixed <- lm(formula = log(velocity_tide128) ~ log(p_tide128) + log(p_tide64) +
                                           log(p_wisk64) + factor(week) + factor(store), data = detergent_DF_2)
reg_tide64_time_nonpromoted_fixed <- lm(formula = log(velocity_tide64) ~ log(p_tide128) + log(p_tide64) +
                                           log(p_wisk64) + factor(week) + factor(store), data = detergent_DF_2)

lm_DF_tide128 <- tidy(reg_tide128_time_nonpromoted_fixed)
lm_DF_tide64 <- tidy(reg_tide64_time_nonpromoted_fixed)

kable(lm_DF_tide128[1:4,], digits = 4)
kable(lm_DF_tide64[1:4,], digits = 4)
```
Yes, the estimates of own and cross price elasticities reveal an improvement over the model specification
in question 4. 
For the regression results of Tide 128 oz, we have the following:
1. The magnitude of own price elasticity decreases more, being -2.3836. It is less extreme.
2. The cross price elasticity of Tide 64 oz on Tide 128 oz is 0.2097 having a positive effect on the demand. That means when the price of Tide 64 oz increases, the demand of Tide 128 oz increases which makes sense (a Tide customer will choose a product of the same brand).
3. The cross price elasticity of Wisk 64 oz on Tide 128 oz is 1.1648 having a positive effect on the demand. That means when the price of Wisk 64 oz increases, the demand of Tide 128 oz increases which makes sense (a Wisk customer will choose a premium brand when the price of a Wisk product increases).

For the regression results of Tide 64 oz, we have the following:
1. The magnitude of own price elasticity decreases more, being -1.4867.
2. The cross price elasticity of Tide 128 oz on Tide 64 oz is 0.9028 having a positive effect on the demand. That means when the price of Tide 128 oz increases, the demand of Tide 64 oz increases which makes sense (a Tide customer will choose a product of the same brand).
3. The cross price elasticity of Wisk 64 oz on Tide 64 oz is -0.2809 having a negative effect on the demand. That means when the price of Wisk 64 oz increases, the demand of Tide 64 oz decreases which does not make more sense as it indicates that these products are complementary and not substitutes.

Compare the estimates to a slightly different regression with the log of unit sales, not log of velocity,
as dependent variable
```{r}
reg_tide128_time_nonpromoted_fixed_2 <- lm(formula = log(q_tide128) ~ log(p_tide128) + log(p_tide64) +
                                           log(p_wisk64) + factor(week) + factor(store), data = detergent_DF_2)
reg_tide64_time_nonpromoted_fixed_2 <- lm(formula = log(q_tide64) ~ log(p_tide128) + log(p_tide64) +
                                           log(p_wisk64) + factor(week) + factor(store), data = detergent_DF_2)

lm_DF_tide128_2 <- tidy(reg_tide128_time_nonpromoted_fixed_2)
lm_DF_tide64_2 <- tidy(reg_tide64_time_nonpromoted_fixed_2)

kable(lm_DF_tide128_2[1:4,], digits = 4)
kable(lm_DF_tide64_2[1:4,], digits = 4)
```

We observe that there is no difference between the results of the regression having the log of unit sales as the dependent variable and the results of the regression having the velocity as the dependent variable.
This absence of difference is expected as we incorporate the store fixed effects in our model.

### Question 6: Pricing and Profitability Analysis
Calculate base (regular) prices, using the data for the non-promoted store-weeks
```{r}
base_price_tide128 <- mean(detergent_DF_2$p_tide128)
base_price_tide128
```
```{r}
base_price_tide64 <- mean(detergent_DF_2$p_tide64)
base_price_tide64
```
Calculate the base volume as average yearly chain-level volume sales
```{r}
count_stores <- length(unique(detergent_DF_2$store))

base_volume_tide128 <- count_stores*52*mean(detergent_DF_2$q_tide128)
base_volume_tide128
```
```{r}
base_volume_tide64 <- count_stores*52*mean(detergent_DF_2$q_tide64)
base_volume_tide64
```
Average yearly base total profit for Tide (sum of profits for Tide 64 and Tide 128)
```{r}
retail_margin <- 0.25
marginal_cost <- 0.027

base_profit_tide128 <- base_volume_tide128*(base_price_tide128*(1-retail_margin)-marginal_cost)
base_profit_tide64 <- base_volume_tide64*(base_price_tide64*(1-retail_margin)-marginal_cost)
base_price_tide <- base_profit_tide128 + base_profit_tide64
base_price_tide
```
Calculate the total new expected volume of Tide, i.e. the new volume of the 128 oz and 64 oz products,
from the following price changes:

A simultaneous 5 percent increase in the prices of Tide 128 and Tide 64
```{r}
del_price_128 <- 0.05
del_price_64 <- 0.05

intercept_tide128 <- reg_tide128_time_nonpromoted_fixed_2$coefficients[[1]]
elasticity_tide128 <- reg_tide128_time_nonpromoted_fixed_2$coefficients[[2]]
elasticity_cross_tide128_64 <- reg_tide128_time_nonpromoted_fixed_2$coefficients[[3]]

intercept_tide64 <- reg_tide64_time_nonpromoted_fixed_2$coefficients[[1]]
elasticity_tide64 <- reg_tide64_time_nonpromoted_fixed_2$coefficients[[3]]
elasticity_cross_tide64_128 <- reg_tide64_time_nonpromoted_fixed_2$coefficients[[2]]

demand_tide128_log <- intercept_tide128 + elasticity_tide128*log(-0.05) + elasticity_cross_tide128_64*log(0.05)

```