---
title: "MLRegression"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This code is giving an example how to estimate a regression given by y = a+b*x+eps, where eps~N(0,sigma^2) using Maximum Likelihood. 

As discussed in class, alternatively one can use a package such as the lm package or, even simpler, use the fact that in a regression beta = (x'x)^-1*x'y.

As part of this example we will simulate data, define a loss function and use the optimizer to find the estimates 



## First Step

Simulate data based on the basic Regression Model y=a+b*x+eps, where eps~N(0,sigma^2).  


```{r}

# Clear All Variables & Clear the Screen
rm(list=ls())
cat("\014")


# Step 1: Make some x data (inputs). For simplicity we only have one x in this example
nobs   <- 300
data.x <- rnorm(n=nobs, mean = 2, sd = 2)


# Step 2: Specify the true parameters of the model 
a.true    <- 1   # Intercept
b.true    <- 2   # Slope
sig.true  <- 1   # Standard Deviation (sigma)


# Step 3: Simulate the y (output) data using x and the model parameters (a,b,sigma)
noise  <- rnorm(n = nobs, mean = 0, sd = sig.true)  # Generate noise
data.y <- a.true + data.x * b.true + noise               # Add noise 


# Step 4: Let’s look at the data:
plot(data.y, data.x, main="Scatterplot",xlab="x ", ylab="y ", pch=19)

```



## Second Step

We need to define the Likelihood to be maximized and need to provide a Loss function that we can give to the optimizer
This has to be a function of the parameters we want to estimate, in our case a(=1), b(=2) and sig(=1). 


```{r}

# Define Define the Loss Function
lm.loss <- function(par) {
  
  a.par     <- par[1]  # The current intercept
  b.par     <- par[2]  # The current slope
  err.sigma <- par[3]  # The current error standard deviation
  
  
  # One issue here is that par[3], the estimate for the standard deviation CANNOT be smaller than zero. We need to ensure that any solution of the optimizer that     suggest par[3]<0 is heavily penalzied (i.e., has a very high loss) and thus is not optimal. The following "if" loop ensures that.  
  if(err.sigma < 0) {deviance <- 10000000}

  
  # If the error standard deviation is valid (i.e.; > 0), then calculate the deviance based on the likelihood
  if(err.sigma > 0) {
  
    # Calculate the likelihood of each data point.
    # Here is where you specify the model and how you calculate likelihoods.
    likelihoods <- dnorm(data.y, mean =  a.par + data.x *b.par, sd = err.sigma)

    
    # Now let's convert the vector of likelihoods to a summary deviance score (-2 times sub log-lik)
    # Calculate log-likelihoods of each data point
    log.likelihoods <- log(likelihoods)
 
    
    # Calculate the deviance (-2 times sum of log-likelihoods)
    deviance <- -2 * sum(log.likelihoods)

  }

return(deviance)

}
```



## Third Step

We need to maximize the Likelihood (or minimize the -Likelihood) using an optimizer


```{r}

# The optim function minimizes a given input
parameter.fits <- optim(par = c(0, 0, 10), fn = lm.loss, hessian = T)

# Display the estimates
parameter.fits$par


# Caculate the Confidence Intervals (like a t-stat)
hessian       <- parameter.fits$hessian
hessian.inv   <- solve(hessian)
parameter.se  <- sqrt(diag(hessian.inv))
CI.matrix     <- as.data.frame(matrix(NA, nrow = 3, ncol = 3))


# Format the 
CI.matrix[1,] <- parameter.fits$par
CI.matrix[2,] <- parameter.fits$par - 1.96 * parameter.se
CI.matrix[3,] <- parameter.fits$par + 1.96 * parameter.se
names(CI.matrix) <- c("a", "b", "sigma")
rownames(CI.matrix) <- c("ML", "95% Lower bound", "95% Upper bound")

# Display CI Matrix
CI.matrix

```




## Alternative: Use an R package

Simply call an R package!

```{r}

# Need to set-up data in one dataframe
data.model <- data.frame("y" = data.y, "x" = data.x)


# Run the Regression (includes an INTERCEPT)
lm.model <- lm(y ~ x, data = data.model)

# Display Results
summary(lm.model)


```
